A Transdiciplinary approach to developing a spectral numerical coastal-flow model based on passive macroalgal rafting phenenomea to shed light on ambient and stochastic dispersal and connectivity of nearshore and offshore marine communities in and around the coast of South Africa: Combining ecology 

### 2019-08-21 literature_review.Rmd
sections under waves edited and added to. simple word output kept but template added for later use. pdf output failed. Will attempt to fix.

---
title: "A Transdiciplinary approach to develop a new spectral numerical coastal-flow model based on natural passive macroalgal rafting phenenomea to shed light on ambient and stochastic dispersal and connectivity of nearshore and offshore marine communities around the coast of South Africa: Combining ecology and 3D hyperspectral hydrodynamic modelling."
author: "Ross Coppin"
date: "20/08/2019"
output: 
  word_document: 
    fig_caption: yes
---

@@

Regardless of the location around the world, waves are a feature of any coastline and the marine environment and are important manifestations of energy in the ocean. Waves are not the movement of water particles but is rather the movement and propagation of energy through the ocean. The source of this energy can be formed locally, such as wind-driven waves, and or from distant locations in the ocean, such as storms. This energy is then transferred from deeper water into shallower water where it plays a role in driving beach morphology. Sand and rock are eroded with each passing wave, thereby playing a role in shaping coastlines. Energy that is left-over after these processes is transferred into heat energy, and heats up the sand and rocks on which each wave hits. This energy can also be harnessed in simple ways such as a surfer catching a wave, or more complex ways such as capturing energy from the ocean environment for electricity production. However, waves are stochastic in nature and therefore the energy that propagates through the ocean is not always consistent. Therefore, it is often difficult to quantify and predict wave energy in the marine environment. Despite this fact, waves are gaining more recognition for the role they play in shaping coastlines, beaches and hence the communities and organisms that depend on these systems.

@@

Waves are formed due to the constant interaction between two forces, which are known as the "generating source" and the “restoration force”. The generating force is the force which pushes water from one layer up into the other, for example pushing water up across the boundary of air. This can also occur in the ocean layers. The ocean layers through differences in density, which in turn is partly driven by abiotic parameters such as temperature and salinity, which creates a boundary for wave energy to move along. Examples of such boundaries are along the ocean surface, pycnoclines, and the sea-floor. optional addition of pycnocline details. In other words, waves occur along the boundaries formed in the ocean by various abiotic processes, and therefore density boundaries are essential to the propagation of energy through the ocean.

@@

Waves are often thought of as an elevation of the sea surface from a specific point over a period of time but this is obvisiously not he case. This is known as *surface elevation* and is the instaneous elevation of the sea surface abouve a specific point in a time record (see figure ???). Although surface elevation does not represent a wave, it can be used to create a wave profile. This is achieved by profiling the the surface elevation between two successive *downward zero-crossings* (see figure ??).

@@

>>>>

---
title: "A Transdiciplinary approach to developing a spectral numerical coastal-flow model based on passive macroalgal rafting phenenomea to shed light on ambient and stochastic dispersal and connectivity of nearshore and offshore marine communities in and around the coast of South Africa: Combining ecology and 3D hyperspectral hydrodynamic modelling."
author: "Ross Coppin"
output: 
  word_document: 
    fig_caption: yes
    toc: yes
---

@@

Regardless of the location around the world, waves are a feature of any coastline and the marine environment, and are important manifestations of energy in the ocean. Waves are not the movement of water particles but rather the movement and propagation of energy through the ocean. The source of this energy can be formed locally, such as wind-driven waves, and or from distant locations in the ocean such as storms, know as *swell*. This energy is transferred from deeper water into shallower water where it plays a role in driving complex marine ecosystems as well as shaping the environments that they live in. Energy that is left-over after these processes is transferred into heat energy, and heats up the sand and rocks on which each wave hits. This energy can also be harnessed in simple ways such as a surfer catching a wave, or more complex ways such as capturing energy from the ocean environment for electricity production. However, waves are stochastic in nature and therefore the energy that propagates through the ocean is not always consistent. Therefore, it is often difficult to quantify and predict wave energy in the marine environment. Despite this fact, waves are gaining more recognition for the role they play in shaping coastlines, beaches and hence the communities and organisms that depend on these systems.

@@

Waves are formed due to the constant interaction between two forces, which are known as the "generating force" and the “restoration force”. The generating force is the force which pushes water from one layer up into the other. Layers in the ocean are formed through differences in temperature (thermocline) and salinity (pycnocline) which creates a boundry for wave energy to move along. In other words, waves occur along the boundaries formed in the ocean by various abiotic processes, and therefore density boundaries are essential to the propagation of energy through the ocean.

@@

Waves are often thought of as an elevation of the sea surface from a specific point over a period of time but this is obvisiously not he case. This is known as *surface elevation* and is the instaneous elevation of the sea surface abouve a specific point in a time record (see figure ???). Although surface elevation does not represent a wave, it can be used to create a wave profile. This is achieved by profiling the the surface elevation between two successive *downward zero-crossings* or *upwardward zero-crossings* (see figure ??). Both zero-crossings are symmetrical and essentially the same statistically. However, in practice the downward zero-crossings are preferred as it takes steepness of a wave into account (the front, see figure ??) which is relevant to characterising breaking waves. It should be noted that surface elevation can be negative while a wave profile cannot.

@@

Characterising waves in a wave record requires compromise both in the statistical sense and the practical, as it is a balance between keeping the record short enough to remain stationary and long enough for a reasonable averages to be calculated. The waves are characterised in terms of wave heights and wave periods for individual waves in the record and then averaged over that specific time. For example, a time record of 15-20 minutes is standard when calculating a wave profile.

```{r Hs_Tp_surface_elevation figure, echo=FALSE, warning=FALSE, fig.cap="The definition of wave height and wave period in a time record of the surface elevation (the wave is defined with downwrd zero-crossings."}

Hs_Tp_elevation <- readPNG("figures/Hs_Tp_surface_elevation.png")
grid.raster(Hs_Tp_elevation)
```
